---
title: "Science Communication Survey Data"
subtitle: Initial data analyses by Victoria
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r data & setup, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(ggsignif)
sc<-read.csv(file="~/Desktop/data/scicomm_data_joined.csv") #import joined file

#import functions for confidence intervals 
#functions from: https://community.rstudio.com/t/computing-confidence-intervals-with-dplyr/31868
lower_ci <- function(mean, se, n, conf_level = 0.95){
  lower_ci <- mean - qt(1 - ((1 - conf_level) / 2), n - 1) * se
}
upper_ci <- function(mean, se, n, conf_level = 0.95){
  upper_ci <- mean + qt(1 - ((1 - conf_level) / 2), n - 1) * se
}

#ggplot plot design for all plots 
plot_design<- theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5) )+
  theme_classic()
```

###Sample size: 
Below is the sample size for the joined data of science communication surveys from spring 2019. 
This sample size has filtered out any students who took < 3 min (on either survey) or had student IDs that did not match in the survey, making it difficult to join their survey and grade records.   

```{r sample size, echo=FALSE, message=FALSE, warning=FALSE}
#sample size in discussion sections
sc %>%
  group_by(group,surveycomplete_begin, surveycomplete_final) %>%
  summarize(n=n()) #sample size per group
```
###Science communication identity 
To briefly analyze effects of the treatment (biweekly science communication exercises) on students' identity as science communicators and educators, I created a cumulative score that summed the students' scores from each of the questions included in survey question 2. Each question could be scored 1 - 10, and there were 23 questions, for a maximum score of 230 points. 

Here is the breakdown of student scores: 
```{r scicomm score, echo=FALSE, message=FALSE, warning= FALSE}
#create "scicomm score" for how much students identify with science communication (overall score for Q2)
#23 questions total, 10 pts each, max score = 230 pts 

sc<-sc %>%
   replace(is.na(.), 0) %>% #change to 0 for ease of math 
   mutate(scicommscore_begin = rowSums(.[23:45])) #adds up all scicomm questions from begin (columns 23 - 45)

sc<-sc %>%
   replace(is.na(.), 0) %>% #change to 0 for ease of math 
   mutate(scicommscore_final = rowSums(.[66:88])) #adds up all scicomm questions from end (columsn 66 - 88)

sc[sc == 0] <- NA #get 0s back to NA

sc$scicommscore_change<-sc$scicommscore_final-sc$scicommscore_begin #calculate change score for individuals w/ pre-post data 

#sanity check: sample size of those with individual change data is 47.
###sum(!is.na(sc$scicomm_score_change))  

#science communication overall score change 
sc %>%
  group_by(group) %>%
  summarize(avg_pre = mean(scicommscore_begin, na.rm=TRUE), sd_pre= sd(scicommscore_begin,na.rm=TRUE), 
            avg_post = mean(scicommscore_final,na.rm=TRUE), sd_post = sd(scicommscore_final,na.rm=TRUE), avg_indiv_change = mean(scicommscore_change, na.rm=TRUE), 
            by_group_diff = avg_post- avg_pre)

```
The graph and linear model output below illustrates differences between the groups in pre and post scores. 
Graphs show means with error bars depicting 95% confidence intervals. 
There is a trend, but these differences are not significant. 

```{r scicomm score analysis, echo=FALSE, message=FALSE, warning=FALSE}
#scid dataset = scicomm identity score overall in a small dataset, converted to long form
#in long form, question will be the question name, survey = begin, final, or change between 
scid<-sc %>% 
 select(id, group, final:coursetotal_grade,scientist_begin:tellstory_begin, scientist_final:tellstory_final, scicommscore_begin:scicommscore_change)%>%
  gather(question, value, final:scicommscore_change)%>%
  separate(col = question, into = c("question", "survey"), sep = "_") 
#minor clean up 
scid$survey<-as.factor(scid$survey) #as factor
scid$question<-as.factor(scid$question)
scid$survey<-recode(scid$survey, "grade" = "NA") #gets rid of the "grade" level in survey


#scicomm score overall barplot
scid %>%
  filter(question=="scicommscore", survey!="change")%>%
  group_by(group, survey)%>%
  summarise(smean = mean(value, na.rm = TRUE),
            ssd = sd(value, na.rm = TRUE),
            count = n()) %>%
  mutate(se = ssd / sqrt(count),
         lower_ci = lower_ci(smean, se, count),
         upper_ci = upper_ci(smean, se, count)) %>%
ggplot( aes(x=as.factor(group), y= as.numeric(smean), fill=survey)) +
  geom_bar(stat="identity", position = position_dodge(), color = "black")+
  geom_errorbar(aes(ymax=upper_ci,ymin=lower_ci),position=position_dodge(0.9), width=0.7)+
  ylim(0,230)+
  labs(x="Survey", y="Science communication identity score")+
  scale_fill_manual(labels = c("Pre", "Post"), values = c("gray", "lightblue"))+
  ggtitle("Science communication identity (overall score)")+
  plot_design

lm1<-scid %>%
  filter(question=="scicommscore", survey!="change")

m1<-lm(value~group*survey, data=lm1)
summary(m1) #no significant effect of anything  
```

###Community belongingness score 
To assess responses to question 3 on the survey, which asked about student attitudes as a sense of belongingness in a community, I created a cumulative score of these 6 questions. Because part 1  "I feel lonely" is negative (sense of belongingness decreases as score increases), I coded this question negatively so that this question's score was substracted from the total score. This gives a maximum total score of 50 for th3 6 questions. 
Below are the summary statistics for this overall score:  
```{r attitudes score, echo=FALSE, message=FALSE, warning=FALSE}
#create an "attitudes score" (max score is 50)
#because the "lonely" question is a negative one, I change the sign so a higher answer in that question lowers the score overall.
sc$lonely_begin<- -(sc$lonely_begin)
sc$lonely_final<- -(sc$lonely_final)

sc<-sc %>%
   replace(is.na(.), 0) %>% #change to 0 for ease of math 
   mutate(attitudescore_begin = rowSums(.[46:51])) #adds up all scicomm questions 

sc<-sc %>%
   replace(is.na(.), 0) %>% #change to 0 for ease of math 
   mutate(attitudescore_final = rowSums(.[89:94])) #adds up all scicomm questions 

sc[sc == 0] <- NA #get 0s back to NA

sc$attitudescore_change<-sc$attitudescore_final-sc$attitudescore_begin #calculate change score for individuals

#attitudes before and after activity 
sc %>%
  group_by(group) %>%
  summarize(avg_pre = mean(attitudescore_begin, na.rm=TRUE), sd_pre = sd(attitudescore_begin,na.rm=TRUE), 
            avg_post = mean(attitudescore_final,na.rm=TRUE), sd_post = sd(attitudescore_final,na.rm=TRUE), avg_indiv_change = mean(attitudescore_change, na.rm=TRUE),
            by_group_diff = avg_post- avg_pre)

```

The graphs below show the community belongingness scores and linear model results between groups. 
Barplots show means and errorbars show 95% confidence intervals. 

```{r attitude score analysis, echo=FALSE, message=FALSE, warning=FALSE}
att<-sc %>%
 select(sisid, group, final:lonely_begin:interested_begin, lonely_final:interested_final, attitudescore_begin:attitudescore_change)%>%
  gather(question, value, final:attitudescore_change)%>%
  separate(col = question, into = c("question","survey"), sep = "_") 

#minor cleanup
att$survey<-as.factor(att$survey)
att$question<-as.factor(att$question)
att$value<-as.numeric(att$value)
att$survey<-recode(att$survey, "grade" = "NA") #get rid of grade level in surveys

#barplot
att %>%
  filter(question=="attitudescore", survey!="change")%>%
  group_by(group, survey)%>%
  summarise(smean = mean(value, na.rm = TRUE),
            ssd = sd(value, na.rm = TRUE),
            count = n()) %>%
  mutate(se = ssd / sqrt(count),
         lower_ci = lower_ci(smean, se, count),
         upper_ci = upper_ci(smean, se, count)) %>%
ggplot( aes(x=as.factor(survey), y= as.numeric(smean), fill=group)) +
  geom_bar(stat="identity", position = position_dodge(), color = "black")+
  geom_errorbar(aes(ymax=upper_ci,ymin=lower_ci),position=position_dodge(0.9), width=0.7)+
  ylim(0,50)+
  geom_signif(comparisons=list(c("begin", "final")), annotations="n.s.", 
              y_position = 40, tip_length = 0.05, vjust = 0,     color="black", textsize = 5 )+
  labs(x="Survey", y="Community belongingness score")+
  scale_x_discrete(labels = c("Pre", "Post"))+
  scale_fill_manual(values = c("gray", "lightblue"))+
  ggtitle("Community belongingness (overall score)")+
  plot_design

lm2<-att %>%
  filter(question=="attitudescore", survey!="change")

m2<-lm(value~group*survey, data=lm2)
summary(m2) #no significant effect of anything

```

###Within-individual differences: 
Only a subset of students completed both the pre and post science communication surveys.  (See sample size in table below).  Within this subset, I looked at change in scores. There is a significant change in individual science communication identity within students in the treatment group (compared to the control group).  

```{r within individual differences, echo = FALSE }

#sample size of those individuals w/ both survey completed, by group 
scid%>%
  filter(question=="scicommscore", survey=="change")%>%
  group_by(group)%>%
  drop_na(value)%>%
  summarize(n=n())

#stats
lm3<-scid%>%
  filter(question=="scicommscore", survey=="change")
m3<-lm(value~group, data=lm3)
summary(m3) #significant effect of treatment on individual change (scicommscore)

#dataset for only students who have both pre and post survey data
scid_both<-sc %>% 
 select(sisid, group, final:coursetotal_grade,scientist_begin:tellstory_begin, scientist_final:tellstory_final, scicommscore_begin:scicommscore_change)%>%
  filter(scicommscore_change != "NA")%>% #subset of students that have both surveys
  gather(question, value, final:scicommscore_change)%>%
  separate(col = question, into = c("question","survey"), sep = "_") 
scid_both$survey<-recode(scid_both$survey, "grade"="NA") #not part of pipe 

#barplot: scicomm score for individual change 
scid_both %>%
  filter(question=="scicommscore", survey=="change")%>%
  group_by(group)%>%
  mutate(perc = (value/230)*100) %>%
  summarise(smean = mean(perc, na.rm = TRUE),
            ssd = sd(perc, na.rm = TRUE),
            count = length(which(!is.na(value)))) %>%
  mutate(se = ssd / sqrt(count),
         lower_ci = lower_ci(smean, se, count),
         upper_ci = upper_ci(smean, se, count)) %>%
ggplot( aes(x=as.factor(group), y= as.numeric(smean), fill=group)) +
  geom_bar(stat="identity", position = position_dodge(), color = "black")+
  geom_errorbar(aes(ymax=upper_ci,ymin=smean),position=position_dodge(0.9), width=0.7)+
  geom_signif(comparisons=list(c("Control", "Treatment")), annotations="p =0.01", 
              y_position = 25, tip_length = 0.05, vjust = 0,     color="black", textsize = 5 )+
  geom_text(aes(label = count, y = 1))+
  ylim(0,30)+
  labs(x="Group", y="% Change in overall science communication identity score")+
  #scale_fill_manual(values = c("gray", "lightblue"))+
  ggtitle("Science communication activities significantly increase individuals' sci-comm identity score")+
  theme(legend.position = "none")+
  plot_design

```

###Grade effects: 
When grouped into treatment groups, it appears the lecture course grades between the groups are incredibly similar.  

```{r grade effects, echo=FALSE, message=FALSE, warning= FALSE}
grades<-c("mt1","mt2","mt3","final", "coursetotal")

#summary stats for grade data
grade_data <- scid%>%
  filter(question %in% grades)%>%
  group_by(question, group)%>%
  summarize(mean_grade = mean(value, na.rm = TRUE), 
            sd_grade = sd(value, na.rm = TRUE))
colnames(grade_data)[1]<- "grade"
grade_data

#barplot of grade data 
scid%>%
  filter(question %in% grades)%>%
  group_by(group, question)%>%
  summarise(smean = mean(value, na.rm = TRUE),
            ssd = sd(value, na.rm = TRUE),
            count = n()) %>%
  mutate(se = ssd / sqrt(count),
         lower_ci = lower_ci(smean, se, count),
         upper_ci = upper_ci(smean, se, count)) %>%
ggplot( aes(x=as.factor(question), y= as.numeric(smean), fill=group)) +
  geom_bar(stat="identity", position = position_dodge(), color = "black")+
  geom_errorbar(aes(ymax=upper_ci,ymin=lower_ci),position=position_dodge(0.9), width=0.7)+
  #ylim(0,230)+
  labs(x="Grade", y="Grade (in points)")+
  scale_x_discrete(labels=c("Total Course Grade", "Final Exam", "Midterm 1", "Midterm 2", "Midterm 3"))+
  scale_fill_manual(values = c("gray", "lightblue"))+
  ggtitle("Grade differences between treatment groups")+
  plot_design
```  
