---
title: "Evaluating surveys without experimental group assignments "
author: "Victoria"
date: "4/24/2020"
output: html_document
---
# Goal: To evaluate surveys missing group assignments
Surveys (pre and post) that were completed but could not be joined with the discussion group assignments. 
Why are these students missing groups? 

Possible explanations: 
*   Did not complete 3+ reflection responses and got filtered out
*   Mistyped their own SIS ID on the survey, making it so that the data could not be joined?
*   Not in the discussion section at all - auditing discussion. 
*   (Pre-survey only) Dropped the class before they could take the post-survey and did not recieve a course grade. 

```{r}
#run sci_comm_data_cleanup_joins.Rmd before running this code. 
  #source this in R?
```

```{r evaluate missing students}
#anti join shows who the individuals are that did surveys but do not have grades/group assignments
pre_missing_group<-anti_join(initial2, disc_grades, by = "sis_id") #25
post_missing_group<-anti_join(final2,disc_grades, by = "sis_id") #7

#WHO ARE THESE STUDENTS?
```

```{r pre-survey}
#PRE: 
#how many simply did not complete 3+ reflection responses? (n = 7 of 25)

#returns those present in discussion grades pre-filtering
few_responses_pre<-inner_join(pre_missing_group, disc_grades_all, by = "sis_id") 

few_responses_pre %>% 
  summarise(n = n(), 
            max_response_no = max(response_count))    #7 individuals 

#remove these from missing group now
pre_miss2 <- anti_join(pre_missing_group, few_responses_pre, by = "sis_id") #now 18

#did the students mistype their own student id? in that case, name should appear even if id doesn't match
disc_grades_all$name <-as.character(disc_grades_all$name)
pre_miss2$last_name<-as.character(pre_miss2$last_name)
disc_grades_all<-separate(disc_grades_all, col = name, into = c("last_name", "first_name"), sep = ",")

#compare last names of those missing groups to the last names in discussion

#not in discussion group assignment (n = 13)
pre_miss2$last_name[str_detect(pre_miss2$last_name, str_c(disc_grades_all$last_name, collapse = "|"), negate = T)]

#yes in dicsussion group assignment (n = 5)
pre_miss2$last_name[str_detect(pre_miss2$last_name, str_c(disc_grades_all$last_name, collapse = "|"), negate = F)]

#of the Yes, did their sis_id's not match?
pre_miss2$sis_id[str_detect(pre_miss2$sis_id, str_c(disc_grades_all$sis_id, collapse = "|"), negate = F)] #returns 0
  #indicates sis_id did not match. will need to manually go back and fix for these 5. 

#of the No, were they enrolled in lecture class and have a grade?
grades<-separate(grades, col = name, into = c("last_name", "first_name"), sep = ",")

#create a vector with last names not assigned a discussion group.
not_in_disc<-pre_miss2$last_name[str_detect(pre_miss2$last_name, str_c(disc_grades_all$last_name, collapse = "|"), negate = T)]

#compare with lecture course grades.
#not in discussion, but present in lecture (had grade) --> auditors
not_in_disc[str_detect(not_in_disc, str_c(grades$last_name, collapse = "|"), negate = F)]

#not in discussion, no grade. --> dropped
not_in_disc[str_detect(not_in_disc, str_c(grades$last_name, collapse = "|"), negate = T)]

```

```{r post-survey}
#POST 
#how many simply did not complete 3+ reflection responses?

#returns those present in discussion grades pre-filtering
few_responses_post<-inner_join(post_missing_group, disc_grades_all, by = "sis_id") 

few_responses_post %>% 
  summarise(n = n(), 
            max_response_no = max(response_count))    #5 out of 7 indivs.

#remove these from missing group now
post_miss2 <- anti_join(post_missing_group, few_responses_post, by = "sis_id") #now 2

#did the students mistype their own student id? in that case, name should appear even if id doesn't match
post_miss2$last_name<-as.character(post_miss2$last_name)

#compare last names of those missing groups to the last names in discussion

#not in discussion group assignment (n = 2 of 2)
post_miss2$last_name[str_detect(post_miss2$last_name, str_c(disc_grades_all$last_name, collapse = "|"), negate = T)]
  #one was NA - did not enter a name or sis-id

# were they enrolled in lecture class and have a grade?

#create a vector with last names not assigned a discussion group.
not_in_disc_post<-post_miss2$last_name[str_detect(post_miss2$last_name, str_c(disc_grades_all$last_name, collapse = "|"), negate = T)]

#compare with lecture course grades.

#none were auditors
not_in_disc_post[str_detect(not_in_disc_post, str_c(grades$last_name, collapse = "|"), negate = F)]

#did not recieve a grade in the course but did post survey...??  (n = 1)
not_in_disc_post[str_detect(not_in_disc_post, str_c(grades$last_name, collapse = "|"), negate = T)]

```


